A memory frontier for complex synapses: Supplementary material

### 2013-08-02 NIPS/rebuttal.txt

Reviewer 1:
Regarding the discrepancy between the envelope and the numerical results: the envelope is just an upper bound and we 
do not claim that it is a tight bound. The reason is that equation (18) is not aa complete set of constraints, as 
discussed below it. In the paragraph from line 377 to 400, we discuss this point further. Note that the numerical 
results are not always instructive as the numerical procedures can be prevented from reaching the global maximum by 
local maxima. This is shown by the fact that our hand designed models can outperform the numerical methods at late 
times. In fact, the apparent drop-off of the solid red curve at very late times comes from not allowing small enough 
epsilon. This should be fixed in fig5.

Reviewer 2:
We'll add a derivation of equation 3 to the supplement. We could add a description in the main text along
the lines of: "The factor of p^infinity describes the synapses being in the steady-state distribution before the
memory is encoded. The factor of (M^pot-M^dep) comes from the encoding of the memory at t=0, with w_ideal being +/-1 
in synapses that are potentiated/depotentiated. The factor of exp(rt W^F) describes the subsequent evolution of the
probability distribution, averaged over all sequences of plasticity events and the factor of w indicates the readout 
via the synaptic weight."
I think we referred to all of the original parts of the supplement in the main paper. I'm not sure how we could
incorporate it into the main paper without exceeding the page limit.

Reviewer 3:
line 195: in the introduction, we cited [17], which describes diversity in synaptic structure across the vertebrate 
brain. This could be related to optimisation for different timescales, but anything more than speculation would 
require a better understanding of the relation between structure and function (which is what we intend to begin with 
this work).
line 220:
line 294: agreed.
line 299: The fact that eta_i depends on M^pot/dep is taken into account. In equation (57) of the supplement, the 
term involving c_g comes from this dependence. If the reviewer is concerned that tho order of the eta_i could change 
during the maximisation procedure: note that necessary conditions for a maximum only require that there is no 
infinitesimal perturbation that increases the area. Therefor we need only consider an infinitesimal neighbourhood of 
the model, in which the order will not change.
line 304: Yes. We'll add this phrase to the text.
line 326: In fig2(a,b), we need not have an equal number of w=+/-1 states (we could change the figure to reflect 
this), but those states are not functionally relevant, as shown by their equivalence to fig2c. The model in fig2c, of 
course, has no room for such asymmetry. The models in figs4,5b do need to have equal numbers of +/- states, asymmetry 
would make them worse. For fig4, the effect of the asymmetry would be reduced in the limit as epsilon -> 0, as all
states other than the end states would have very small p^infinity.
line 415: It is only for t_0 > sqrt(N)M that the models that nearly touch the envelope are linear chains. The model 
in fig5b is not a linear chain, as it has shortcut transitions, but they are only the best models (that we have 
found) for times t_0 < sqrt(N)M. Would it help if we repeated fig4 as fig5c?

### 2013-08-02 NIPS/rebuttal.txt

Reviewer 1: We briefly discussed a suggestion for biological
experiments to test the theory in lines 428. We could expand on
this. If we measured pre and post-synaptic spike trains, and also
recorded changes in post-synaptic potentials to measure changes in
synaptic weight, we could use hidden Markov model techniques to find the best-fit Markov

Regarding the discrepancy between the envelope and the
numerical results: the envelope is just an upper bound and we do
not claim that it is a tight bound. The reason is that equation
(18) is not aa complete set of constraints, as discussed below it.
In the paragraph from line 377 to 400, we discuss this point
further. Note that the numerical results are not always instructive
as the numerical procedures can be prevented from reaching the
global maximum by local maxima. This is shown by the fact that our
hand designed models can outperform the numerical methods at late
times. In fact, the apparent drop-off of the solid red curve at
very late times comes from not allowing small enough epsilon. This
should be fixed in fig5.

Reviewer 2: We'll add a derivation of equation 3 to the supplement.
We could add a description in the main text along the lines of:
"The factor of p^infinity describes the synapses being in the
steady-state distribution before the memory is encoded. The factor
of (M^pot-M^dep) comes from the encoding of the memory at t=0, with
w_ideal being +/-1 in synapses that are potentiated/depotentiated.
The factor of exp(rt W^F) describes the subsequent evolution of the
probability distribution, averaged over all sequences of plasticity
events and the factor of w indicates the readout via the synaptic
weight."
I think we referred to all of the original parts of the supplement
in the main paper. I'm not sure how we could incorporate more into
the main paper without exceeding the page limit.

### 2013-08-02 NIPS/rebuttal.txt

Reviewer 1: We briefly discussed a suggestion for biological
experiments to test the theory in line 428. We could expand on
this. If we measured pre and post-synaptic spike trains, and also
recorded changes in post-synaptic potentials to measure changes in
synaptic weight, we could use hidden Markov model techniques to
find the best-fit synaptic model. However, as shown in this paper,
which model is best depends on the range of time-scales of
interest, so any predictions would also require understanding of
what timescales the particular synapse needs to operate on.
Regarding the discrepancy between the envelope and the
numerical results: the envelope is just an upper bound and we do
not claim that it is a tight bound. The reason is that equation
(18) is not aa complete set of constraints, as discussed below it.
In the paragraph from line 377 to 400, we discuss this point
further. Note that the numerical results are not always instructive
as the numerical procedures can be prevented from reaching the
global maximum by local maxima. This is shown by the fact that our
hand designed models can outperform the numerical methods at late
times. In fact, the apparent drop-off of the solid red curve at
very late times comes from not allowing small enough epsilon. This
should be fixed in fig5.
Yes, the level of noise tolerance is an important consideration.
Note that the stochasticity of the models is an expression of
biological noise. In fact, the type of noise described by the
reviewer could be included by adding extra states to the model with
the different transition rates. However, the synapses should not be
allowed to optimise all of this noise away. It would be interesting
to see how our results change if we put lower limits on the noise
levels. I think this would be beyond the scope of this work, but
should be added to the discussion section.

@@

Reviewer 2: We'll add a derivation of equation 3 to the supplement.
We could add a description in the main text along the lines of:
"The factor of p^infinity describes the synapses being in the
steady-state distribution before the memory is encoded. The factor
of (M^pot-M^dep) comes from the encoding of the memory at t=0, with
w_ideal being +/-1 in synapses that are potentiated/depotentiated.
The factor of exp(rt W^F) describes the subsequent evolution of the
probability distribution, averaged over all sequences of plasticity
events and the factor of w indicates the readout via the synaptic
weight."
I think we referred to all of the original parts of the supplement
in the main paper. I'm not sure how we could incorporate more into
the main paper without exceeding the page limit.

Reviewer 3:
line 195: in the introduction, we cited [17], which describes
diversity in synaptic structure across the vertebrate brain. This
could be related to optimisation for different timescales, but
anything more than speculation would require a better understanding
of the relation between structure and function (which is what we
intend to begin with this work).
line 220: Yes, fig2a,b is an example of why it is difficult to map
molecular states to functional states as, despite appearances,
these models actually only have two functional states due to their
equivalence to fig 2c. One experimental investigation of this could
be along the lines of the experiment described in line 428 (see
first paragraph of our reply to reviewer 1 above for elaboration).
Presumably we will find fewer functional states than molecular
states. Making the link between molecular and functional states
will, unfortunately, be much harder
line 294: agreed.
line 299: The fact that eta_i depends on M^pot/dep is taken into
account. In equation (57) of the supplement, the term involving c_g
comes from this dependence. If the reviewer is concerned that tho
order of the eta_i could change during the maximisation procedure:
note that necessary conditions for a maximum only require that
there is no infinitesimal perturbation that increases the area.
Therefor we need only consider an infinitesimal neighbourhood of
the model, in which the order will not change.
line 304: Yes. We'll add this phrase to the text.
line 326: In fig2(a,b), we need not have an equal number of w=+/-1
states (we could change the figure to reflect this), but those
states are not functionally relevant, as shown by their equivalence
to fig2c. The model in fig2c, of course, has no room for such
asymmetry. The models in figs4,5b do need to have equal numbers of
+/- states, asymmetry would make them worse. For fig4, the effect
of the asymmetry would be reduced in the limit as epsilon -> 0, as
all states other than the end states would have very small
p^infinity.
line 415: It is only for t_0 > sqrt(N)M that the models that nearly
touch the envelope are linear chains. The model in fig5b is not a
linear chain, as it has shortcut transitions, but they are only the
best models (that we have found) for times t_0 < sqrt(N)M. Would it
help if we repeated fig4 as fig5c?

### 2013-08-05 NIPS/rebuttal.txt

Reviewer 1:
We would like to link our theory to experiment, but we don't think
that the necessary experiments have been performed yet. We briefly
discussed a suggestion for biological experiments to test the
theory in line 428. We could expand on this. If we measured pre and
post-synaptic spike trains, and also recorded changes in
post-synaptic potentials/currents from a single synapse to measure
changes in synaptic weight, we could use hidden Markov model
techniques to find the best-fit synaptic model. However, as shown
in this paper, which model is best depends on the range of
time-scales of interest, so any predictions would also require
understanding of what timescales the particular synapse needs to
operate on. We'd also need to cut something else to stay within the
page limit.
Regarding the discrepancy between the envelope and the
numerical results: the envelope is just an upper bound and we do
not claim that it is a tight bound. The reason is that equation
(18) is not aa complete set of constraints, as discussed below it.
In the paragraph from line 377 to 400, we discuss this point
further. Note that the numerical results are not always instructive
as the numerical procedures can be prevented from reaching the
global maximum by local maxima. This is shown by the fact that our
hand designed models can outperform the numerical methods at late
times. In fact, the apparent drop-off of the solid red curve at
very late times comes from not allowing small enough epsilon. This
should be fixed in fig5, showing the solid red curve hugging the
green one at late times.
We agree that the level of noise tolerance is an important
consideration. Note that the stochasticity of the models is an
expression of biological noise. In fact, the type of noise
described by the reviewer could be included by adding extra states
to the model with the different transition rates. However, the
synapses should not be allowed to optimise this noise away.
It would be interesting to see how our results change if we put
lower limits on the noise levels, or incorporated noise in the
manner suggested by the reviewer. I think this would be beyond the
scope of this work, but could be mentioned in the discussion section.

Reviewer 2:
We'll add a derivation of equation 3 to the supplement.
We could add a description in the main text along the lines of:
"The factor of p^infinity describes the synapses being in the
steady-state distribution before the memory is encoded. The factor
of (M^pot-M^dep) comes from the encoding of the memory at t=0, with
w_ideal being +/-1 in synapses that are potentiated/depotentiated.
The factor of exp(rt W^F) describes the subsequent evolution of the
probability distribution, averaged over all sequences of plasticity
events and the factor of w indicates the readout via the synaptic
weight."
I think we referred to all of the original parts of the supplement
in the main paper. I'm not sure how we could incorporate more into
the main paper without exceeding the page limit. Section 1 of the
supplement consists of simple generalisations of discrete time
results to continuous time, so it doesn't really merit inclusion in
the main text. We put references at the start of the section. Would
it make it clearer if we added these references throughout the section?

Reviewer 3:
line 195: in the introduction, we cited [17], which describes
diversity in synaptic structure across the vertebrate brain. This
could be related to optimisation for different timescales, but
anything more than speculation would require a better understanding
of the relation between structure and function (which is what we
intend to begin with this work).
line 220: Yes, fig2a,b is an example of why it is difficult to map
molecular states to functional states as, despite appearances,
these models actually only have two functional states due to their
equivalence to fig 2c. One experimental investigation of this could
be along the lines of the experiment described in line 428 (see
first paragraph of our reply to reviewer 1 above for elaboration).
Presumably we will find fewer functional states than molecular
states. Making the link between molecular and functional states
will, unfortunately, be much harder.
line 294: agreed.
line 299: The fact that eta_i depends on M^pot/dep is taken into
account. In equation (57) of the supplement, the term involving c_g
comes from this dependence. If the reviewer is concerned that the
order of the eta_i could change during the maximisation procedure:
note that necessary conditions for a maximum only require that
there is no infinitesimal perturbation that increases the area.
Therefor we need only consider an infinitesimal neighbourhood of
the model, in which the order will not change. Should this be
mentioned in section 3.2?
line 304: Yes. We'll add this phrase to the text.
line 326: In fig2(a,b), we need not have an equal number of w=+/-1
states (we could change the figure to reflect this), but those
states are not functionally relevant, as shown by their equivalence
to fig2c. The model in fig2c, of course, has no room for such
asymmetry. The models in figs4,5b do need to have equal numbers of
+/- states, asymmetry would make them worse. For fig4, the effect
of the asymmetry would go away in the limit as epsilon -> 0, as
all states other than the end states would have very small
p^infinity. Should this be mentioned in section 4?
line 415: It is only for t_0 > sqrt(N)M that the models that nearly
touch the envelope are linear chains. The models in fig5b is not a
linear chain, as it has shortcut transitions, but these are only the
best models (that we have found) for times t_0 < sqrt(N)M. Would it
help if we repeated fig4 as fig5c?

### 2013-08-16 NIPS/SynapseNIPSsupp.tex
fixed bookmark in supp
working on teo-time env
updated NIPS supp

\begin{defn}[Fundamental matrix]
  %
  \begin{equation}\label{eq:funddef}
    \fund \equiv (-\MM + \onev\pib)\inv,
  \end{equation}
  %
  where $\pib$ is any row-vector with $\pib\onev=1/\tau\neq0$.
\end{defn}
Note that the canonical choice for the discrete time version, $\pib=\eq$, is not available here due to problems with units.
It will be helpful to choose $\pib$ to be independent of $\MM$, \eg $\pib=\onev\trans/(n\tau)$.
All quantities that we calculate using $\fund$ below will be independent of this choice.

@@

The off-diagonal mean first passage times are given by
%
\begin{equation}\label{eq:fptfund}
  \fptb = (\onem\fund\dg - \fund)\D.
\end{equation}
%
or in component form of \eqref{eq:fptfund}:
%
\begin{equation}\label{eq:fptfundcmpt}
  \fptb_{ij} = \frac{\fund_{jj}-\fund_{ij}}{\eq_j}.
\end{equation}
%

@@

\begin{thm}\label{th:kemenyconst}
  The quantity
  %
  \begin{equation}\label{eq:mixdef}
    \eta \equiv \sum_j \fptb_{ij}\eq_j
  \end{equation}
  %
  is independent of $i$.
\end{thm}
\begin{proof}
  We use \eqref{eq:fptfund}, \eqref{eq:fundrowsum} and the transpose of \eqref{eq:pdotD}:
  %
  \begin{equation*}
    \begin{aligned}
      \fptb (\eq)\trans &= (\onem\fund\dg - \fund) \D (\eq)\trans \\
        &= (\onev\onev\trans\fund\dg - \fund) \onev \\
        &= (\onev\trans\fund\dg\onev)\onev - \fund\onev \\
        &= (\tr\fund - \tau)\onev.
    \end{aligned}
  \end{equation*}
  %
  which proves \eqref{eq:mixdef} with $\eta = \tr\fund - \tau$.
\end{proof}

Note that it is essential that we use $\fptb$ and not $\fpt$ here, as that would lead to $\eta_i=\eta+\Lb_{ii}$, unlike the discrete time version \cite{hunter2006mixing} where this would only shift $\eta$ by 1.

@@

As discussed in the main text, the signal-to-noise ratio is given by
%
\begin{equation}\label{eq:SNRcurve}
  \SNR(t) = \sqrt{N} (2f\pot f\dep ) \eq (\W\pot  - \W\dep )\e^{rt\frg}\w.
\end{equation}
%

@@

\subsection{Lumpability}\label{sec:SNRlump}

@@

Using $\eq\frg=0$, we can write the initial SNR as
%
\begin{equation}\label{eq:init}
  \frac{\SNR(0)}{\sqrt{N}} = I = ({\eq}^++{\eq}^-)(f\pot \W\pot -f\dep \W\dep )(\onev^+-\onev^-).
\end{equation}
%
Using $\W\potdep (\onev^++\onev^-)=0$ and \eqref{eq:compflux}:
%
\begin{equation*}
  r{\eq}^-(f\pot \W\pot +f\dep \W\dep )\onev^+ = \F_{-+} = \F_{+-} = r{\eq}^+(f\pot \W\pot +f\dep \W\dep )\onev^-,
\end{equation*}
%
we can rewrite \eqref{eq:init} as
%
\begin{equation}\label{eq:initflux}
  I = \frac{4\F_{-+}}{r} - 4{\eq}^+f\pot \W\pot \onev^- - 4{\eq}^-f\dep \W\dep \onev^+.
\end{equation}
%
The last two terms are guaranteed to be negative, as the diagonal parts of $\W\potdep $ cannot contribute.
Therefore
%
\begin{equation}\label{eq:initfluxineq}
  \SNR(0) \leq \frac{4\sqrt{N}\F_{-+}}{r}.
\end{equation}
%
This inequality is saturated if potentiation never takes it from a $+$ state to a $-$ state and depression never takes it from a $-$ state to a $+$ state.

@@

\subsection{Derivatives \wrt \texorpdfstring{$\W\potdep $}{W+/-}}\label{sec:deriv}

>>>>

\begin{defn}[Fundamental matrix]
  For discrete time, the generalized fundamental matrix was defined in \cite{Kemeny1981const}.
  For continuous time, we define:
  %
  \begin{equation}\label{eq:funddef}
    \fund \equiv (-\MM + \onev\pib)\inv,
  \end{equation}
  %
  where $\pib$ is any row-vector with $\pib\onev=1/\tau\neq0$.
\end{defn}
Note that the canonical choice for the discrete time version, $\pib=\eq$, is not available here due to problems with units.
It will be helpful to choose $\pib$ to be independent of $\MM$, \eg $\pib=\onev\trans/(n\tau)$.
All quantities that we calculate using $\fund$ below will be independent of this choice.

@@

The off-diagonal mean first passage times are given by
%
\begin{equation}\label{eq:fptfund}
  \fptb = (\onem\fund\dg - \fund)\D.
\end{equation}
%
or in component form:
%
\begin{equation}\label{eq:fptfundcmpt}
  \fptb_{ij} = \frac{\fund_{jj}-\fund_{ij}}{\eq_j}.
\end{equation}
%

@@

\begin{thm} \mbox{}\label{th:kemenyconst}
  The quantity
  %
  \begin{equation}\label{eq:mixdef}
    \eta \equiv \sum_j \fptb_{ij}\eq_j
  \end{equation}
  %
  is independent of $i$.
\end{thm}
\begin{proof}
  For discrete time, a proof can be found in \cite[Th.4.4.10]{kemeny1960finite}.
  For continuous time, we use \eqref{eq:fptfund}, \eqref{eq:fundrowsum} and the transpose of \eqref{eq:pdotD}:
  %
  \begin{equation*}
    \begin{aligned}
      \fptb (\eq)\trans &= (\onem\fund\dg - \fund) \D (\eq)\trans \\
        &= (\onev\onev\trans\fund\dg - \fund) \onev \\
        &= (\onev\trans\fund\dg\onev)\onev - \fund\onev \\
        &= (\tr\fund - \tau)\onev.
    \end{aligned}
  \end{equation*}
  %
  which proves \eqref{eq:mixdef} with $\eta = \tr\fund - \tau$.
\end{proof}

Note that it is essential that we use $\fptb$ and not $\fpt$ here, as that would lead to $\eta_i=\eta+\Lb_{ii}$, unlike the discrete time version, where this would only shift $\eta$ by 1.

@@

As discussed in the main text, the signal-to-noise ratio is given by
%
\begin{equation}\label{eq:SNRdef}
  \snr(t) = \frac{\av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}}
     {\sqrt{\var(\wv_\text{ideal}\cdt\wv(\infty))}}.
\end{equation}
%
First, let's look at the denominator, remembering that the states and plasticity events of each synapse are independent and identically distributed:
%
\begin{equation}\label{eq:noise}
\begin{aligned}
  \var(\wv_\text{ideal}\cdt\wv(\infty))
    &= \sum_{\alpha\beta} \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty) \wv^\beta_\text{ideal}\wv^\beta(\infty)}
    - \prn{\sum_\alpha \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}}^2 \\
    &= \sum_{\alpha} \av{(\wv^\alpha_\text{ideal})^2(\wv^\alpha(\infty))^2}
    + \sum_{\alpha\neq\beta} \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}\av{\smash{\wv^\beta}_\text{ideal}\smash{\wv^\beta}(\infty)}
    \\&\phantom{= \sum_{\alpha}  \av{(\wv^\alpha_\text{ideal})^2(\wv^\alpha(\infty))^2}}
    - \prn{\sum_\alpha \av{\wv^\alpha_\text{ideal}\wv^\alpha(\infty)}}^2  \\
    &= N\av{1}
    + N(N-1)\av{\wv^1_\text{ideal}\wv^1(\infty)}^2
    - N^2\av{\wv^1_\text{ideal}\wv^1(\infty)}^2 \\
    &= N(1-\av{\wv^1_\text{ideal}\wv^1(\infty)}^2),
\end{aligned}
\end{equation}
%
where we used $\wv^\alpha=\pm1$.

For the numerator, we can write
%
\begin{equation}\label{eq:overlapex}
  \av{\wv_\text{ideal}\cdt\wv(t)} = \sum_\alpha\av{\wv^\alpha_\text{ideal}\wv^\alpha(t)}
   = N \av{\wv^1_\text{ideal}\wv^1(t)},
\end{equation}
%
Noting that $\wvi=\pm1$ with probability $f\potdep$,
%
\begin{equation}\label{eq:overlapstate}
\begin{aligned}
  \av{\wv^1_\text{ideal}\wv^1(t)} &= f\pot \av{\wv^1(t)}_{\text{pot},t=0} - f\dep \av{\wv^1(t)}_{\text{dep},t=0} \\
   &= f\pot \sum_i P(\text{state}=i,t \mid \text{pot},0) \w_i - f\dep \sum_i P(\text{state}=i,t \mid \text{dep},0) \w_i.
\end{aligned}
\end{equation}
%
From the previous section,
%
\begin{equation}\label{eq:stateprob}
  P(\text{state}=i,t \mid \text{pot/dep},0) = \brk{ \eq \M\potdep\, \e^{rt\frg} }_i,
\end{equation}
%
which describes the synapses starting in the equilibrium distribution, changing state due to the plasticity event at $t=0$ and subsequent evolution according to \eqref{eq:rowsum} due to plasticity events uncorrelated with the special event at $t=0$.%
\footnote{Note that expanding the exponential gives
%
\begin{equation*}
  \e^{rt\frg} =  \sum_{n=0}^\infty \frac{(rt)^n\,\e^{-rt}}{n!} \sum_{m=0}^n (f\pot)^m (f\dep)^{n-m}
  \brk{ \M\pot \M\dep \M\pot \M\pot \ldots + \text{permutations}}.
\end{equation*}
%
Thus, evolving according to \eqref{eq:rowsum} results in averaging over all sequences of plasticity events,
as we only need linear expectations of $\wv(t)$ in the end.}
This results in
%
\begin{equation}\label{eq:overlapwt}
\begin{aligned}
  \av{\wv^1_\text{ideal}\wv^1(t)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \e^{rt\frg} \w, \\
  \av{\wv^1_\text{ideal}\wv^1(\infty)} &= \eq (f\pot\M\pot-f\dep\M\dep)\, \onev\eq \w \\
         &= \eq (f\pot\onev-f\dep\onev)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \w \\
         &=  (f\pot-f\dep)\, \eq \e^{rt\frg} \w .
\end{aligned}
\end{equation}
%
Combining these allows us to write the numerator as
%
\begin{equation}\label{eq:signal}
\begin{aligned}
  \av{\wv_\text{ideal}\cdt\wv(t)} - \av{\wv_\text{ideal}\cdt\wv(\infty)}
    &= N \eq (f\pot(\M\pot-\I)-f\dep(\M\dep-\I))\, \e^{rt\frg} \w \\
    &= N \eq (f\pot(\W\pot-\frg)-f\dep(\W\dep-\frg))\, \e^{rt\frg} \w \\
%    &= N \eq (f\pot(f\dep\W\pot-f\dep\W\dep)-f\dep(f\pot\W\dep-f\pot\W\pot))\,\e^{rt\frg}\w \\
    &= N (2f\pot f\dep)\eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w.
\end{aligned}
\end{equation}
%
where we used $\eq\frg=0$ in going from the first to second lines.
Combining with \eqref{eq:noise} gives
%
\begin{equation}\label{eq:SNRcurveExact}
  \SNR(t) = \frac{\sqrt{N} (2f\pot f\dep ) \eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w}
                 {\sqrt{1-(f\pot-f\dep)^2(\eq_+-\eq_-)^2}}.
\end{equation}
%
The denominator will not play any role in what follows, as the models that maximize the various measures of memory performance all have some sort of balance between potentiation and depression, either with $f\pot=f\dep$ or $\eq_+=\eq_-$.
We can set the denominator to 1 without changing any of our results.

Which results in our final formula:
%
\begin{equation}\label{eq:SNRcurve}
  \SNR(t) = \sqrt{N} (2f\pot f\dep )\, \eq (\W\pot  - \W\dep )\,\e^{rt\frg}\w.
\end{equation}
%

@@

\subsection{Lumpability and the SNR curve}\label{sec:SNRlump}

@@

Using $\eq\frg=0$ and the first line of \eqref{eq:signal}, we can write the initial SNR as
%
\begin{equation}\label{eq:init}
  \frac{\SNR(0)}{\sqrt{N}} = I = ({\eq}^++{\eq}^-)(f\pot \W\pot -f\dep \W\dep )(\onev^+-\onev^-).
\end{equation}
%
Using $\W\potdep (\onev^++\onev^-)=0$ and \eqref{eq:compflux}:
%
\begin{equation*}
  r{\eq}^-(f\pot \W\pot +f\dep \W\dep )\onev^+ = \F_{-+} = \F_{+-} = r{\eq}^+(f\pot \W\pot +f\dep \W\dep )\onev^-,
\end{equation*}
%
we can rewrite \eqref{eq:init} as
%
\begin{equation}\label{eq:initflux}
  I = \frac{4\F_{-+}}{r} - 4{\eq}^+f\pot \W\pot \onev^- - 4{\eq}^-f\dep \W\dep \onev^+.
\end{equation}
%
The last two terms are guaranteed to be negative, as the diagonal parts of $\W\potdep $ cannot contribute.
Therefore
%
\begin{equation}\label{eq:initfluxineq}
  \SNR(0) \leq \frac{4\sqrt{N}\F_{-+}}{r}.
\end{equation}
%
This inequality is saturated if potentiation never takes it from a $+$ state to a $-$ state and depression never takes it from a $-$ state to a $+$ state.

@@

\subsection{Derivatives \wrt \texorpdfstring{$\W\potdep $}{W(pot/dep)}}\label{sec:deriv}
